---
title: "PHS_Trends"
author: "Awoe"
date: "4/21/2021"
output: html_document
---

Goal: Prepare optum raw health hospitalizations data to merge with temperature data.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(tigris)
library(leaflet)
library(sp)
library(rgeos)
library(dplyr)
library(mapview)
library(ggplot2)


# install.packages("censusapi")
library(censusapi)

memory.limit(64000)
```


Step 1: Initializing -- Pulling Optum Data; we're going to focus on cleaning the monthly data set, but you can pull yearly or seasonly. 
```{r pulling-data}
data<- readRDS("G:/Shared drives/SFBI-Restricted/PHS/optum_summary_year.rds")
#data_month <- readRDS("G:/Shared drives/SFBI-Restricted/PHS/optum_summary_month.rds")
#data_season <- readRDS("G:/Shared drives/SFBI-Restricted/PHS/optum_summary_season.rds")
```

Step 2: Pulling Zipcode and Population Data
```{r setting up for creating our sf later}
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs" 
# bay_county_names <-
#   c(
#     "Alameda", #zip codes: 94501, 94502, 94606
#     "Contra Costa", #start with 94
#     "Marin", #start with 94
#     "Napa", #start with 94
#     "San Francisco", #start with 94
#     "San Mateo", #starts with 94
#     "Santa Clara", #start with 94 and 95
#     "Solano", #start with 94 and 95
#     "Sonoma" #start with 94 and 95
#   )

CA_counties <-
  counties("CA", cb = T, progress_bar = F) 

#grabbing the ZCTAS from tigris
zipcodes<- zctas(cb = TRUE, starts_with = c("94","95","90", "91","93","92","96"))

```

Step 3: Three goals here: 
a. our data set includes more than just california so here, we will filter zipcodes for just CA
b. we're going to filter for just heat-related hospitalizations in via ICD codes in the DIAG column
c. create shapefiles from the zipcodes so we can map them
```{r filtering just for heat hospitalizations across CA | 3ab}
### hiding code that filters for bay area. Just gonna plot entire california; assuming 
### that all the counties are in CA. 

#YEARLY DATA~~~~~~~~~~~~~~~~~~

#filter the data within California's ZCTAs based on the first two characters/"digits" of the data's ZIPCODE_5 
data_CA <- data %>%
 filter(substr(ZIPCODE_5,1,2) %in% c("90","91","92","93","94","95","96"))  %>%  
  filter(
    substr(DIAG, 1, 3) %in% c(
      "992",
      "E87",
      "T67"
    ) | substr(DIAG, 1, 4) %in% c(
      "2760",
      "9944",
      "9945",
      "T732",
      "T733"
    )
  )

#MONTHLY DATA~~~~~~~~~~~~~~~~~~~~
# data_CA <- data_month %>%
#  filter(substr(ZIPCODE_5,1,2) %in% c("94","95","90", "91","93","92","96"))  %>%  
#   filter(
#     substr(DIAG, 1, 3) %in% c(
#       "992",
#       "E87",
#       "T67"
#     ) | substr(DIAG, 1, 4) %in% c(
#       "2760",
#       "9944",
#       "9945",
#       "T732",
#       "T733"
#     )
#   )
```

Step 3c: 
```{r assigning geometries to zipcodes | 3c}

#data_CA has multiple observations for each unique 1470 ZCTAs making up our 5mill observations. So we'll look at the unique ones then fill in later with a left_join().
data_CA_unique <- data_CA %>% 
  ungroup() %>% 
  #select(-c("year","DIAG","n")) %>% 
  mutate(geometry = NA) %>% #creating an empty columns for the geometry
  distinct(ZIPCODE_5,.keep_all = TRUE) #grabbing just the unique/distinct ZCTAs
 
#loop that aggregates zipcodes to match the optum dataset
for (i in 1:nrow(data_CA_unique)){
  #grab the first z1
  zcta <- str_split(data_CA_unique$ZIPCODE_5[i], "_") %>% unlist() #split the zip codes
  zcta_geom <- zipcodes %>% filter(ZCTA5CE10 %in% zcta) #grabbing the geometry of the zctas from the bay area ZCTAs
  zcta_pop <- CA_pop_ages %>% filter(zip_code_tabulation_area %in% zcta) #grabbing zipcodes from our census data to preserve vector length
  if(nrow(zcta_geom) > 0){ #if there is more than 0 in the dataframe then
    #grab the geometry
    zcta_geom <- zcta_geom %>% st_union() %>% st_cast("MULTIPOLYGON") #union/join the geometries together and make it a multipolygon
    data_CA_unique$geometry[i] <- zcta_geom #assign it to the geometry column
  }
}

# Leftjoining our unique dataframe with our full dataframe ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#YEARLY
data_CA_geom_pop <- left_join(data_CA %>% ungroup(), data_CA_unique, by = c("ZIPCODE_5"))#combining unique geometries with the entire data set

#MONTHLY leftjoin

# data_CA_month_geom_pop <- left_join(data_CA %>% ungroup(), data_CA_unique, by = c("ZIPCODE_5","year","DIAG","n","month"))#combining unique geometries with the entire data set

# create our SF object~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#YEARLY 
#final <- data_CA_geom %>% 
data_CA_geom %>% 
  st_as_sf() %>% #make the data a sf object
  st_set_crs(4269) %>% #assign it a coordinate reference system (crs)
  .[CA_counties %>% st_transform(projection) %>% st_buffer(-100) %>% st_transform(4269), ]
  #only keep the ZCTAs that are within the Bay Area by creating an inwards buffer of the Bay Area counties' geometries
# could comment the bay_counties portion two lines above if doing all of CA

#MONTHLY~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# #final <- data_CA_geom %>% 
# data_CA_month_geom_pop %>% 
#   st_as_sf() %>% #make the data a sf object
#   st_set_crs(4269) %>% #assign it a coordinate reference system (crs)
#   .[CA_counties %>% st_transform(projection) %>% st_buffer(-100) %>% st_transform(4269), ]
#   #only keep the ZCTAs that are within the Bay Area by creating an inwards buffer of the Bay Area counties' geometries
# # could comment the bay_counties portion two lines above if doing all of CA
```


Step 4: Saving Data -- this will be used to combine with heat data and population added in HeatnHealth_CA
```{r saving data}

setwd("G:/Shared drives/SFBI-Restricted/PHS")

# save data
saveRDS(data_CA_geom, "CA_year_ZCTAS.rds")
# saveRDS(data_CA_month_geom_pop, "CA_month_ZCTAS.rds")
# saveRDS(final, "bay_area_season_ZCTAS.rds")



```



